\\ Preventing Leakage of Distribution in Order Preserving Encryption Schemes \\ \vspace*{.3in} \small Hrishikesh Joshi, Max Kolysh, Mari Miyachi, John Wang \\ \small 6.857 Final Paper \\ \small May 15, 2013

### 2013-05-15 paper.tex
proofing

Order-preserving encryption schemes serve to generate ciphertexts that preserve the numerical ordering of the plaintexts. Such schemes provide an entity storing encrypted data, such as a mail server or a server interacting with a database, the ability to perform order comparison operations such as sorting, range query checks, and ranking, among other applications. The ideal security goal for an order-preserving encryption scheme, IND-OCPA, is for the scheme to reveal nothing except for the order of the plaintexts. \\ \\
For more than a decade, researchers have invented such schemes with varying degrees of leakage of plaintexts and other attributes of the data. Recently, the first order-preserving encryption scheme that satisfies IND-OCPA, mutable order-preserving encoding (mOPE), was proposed by Popa et al (2013). mOPE works by creating a balanced search tree containing all the encrypted data on a server, giving the server the ability the order the tree based on the plaintexts without giving the server information regarding the plaintexts themselves. All encrypted values are represented by a node in the balanced search tree that has an encoding, the Order Preserving Encoding (OPE), associated with it which represents its path on the tree from the root to the node, allowing the server the ability to maintain the order of the plaintexts. \\ \\ 
One disadvantage of mOPE, however, is that it can leak information about the distribution of plaintext values since multiple entries of the same plaintext value are maintainted in the same node in the tree with a counter. In this paper, we present two schemes that address this issue: Distribution Confidentiality Scheme (DCS) and Random Tiebreaking Order Preservering Encryption (RTOPE). Further, we define a new security guarantee, Indistinguisability Under Chosen Plaintext Distribution Attack (IND-CPDA), which these schemes are intended to satisfy. \\ \\

@@

The mOPE scheme by Popa et al (2013) is the first scheme to achieve ideal IND-OCPA security, and our schemes are built upon mOPE. Prior work on range query schemes have provided schemes that leak information about the underlying plaintext. Shi et al (2007) designed an encryption scheme that leaks the decrypted data for values that fall within the range query, but preserves the privacy of all other values. We chose to build on schemes on existing OPE schemes, instead of range query schemes, to provide the security of the plaintext even within the range queries. \\ \\
Further, Popa et al (2013) provides a new security definition, same-time OPE security, which dictates that an adversary can learn the order only for values stored in the application at the same time, and is stronger than IND-OCPA. Our security guarantee of IND-CPDA, explained later in the paper, asks that the ciphertexts not leak the distribution of values. 

@@

In the basic version of the scheme, let us assume that we want to encrypt some plaintext value $x \in \mathrm{Z}$ and send it to the server like in Popa et al (2013)'s mOPE scheme. For DCS, there is an additional parameter $\rho$ which will govern the probability with which the server sees a repeated value. We will generate a random value $r \in [0,1)$ which contains $\rho$ bits, denoted $r_{\rho}$, and append it to the end of $x$. The resulting value $x' = x + r_{\rho}$ will be encrypted with DET. Popa et al (2013)'s mOPE scheme will work the same as before, but now acting on $x'$. \\ \\
Inserting in this scheme is simple, since nothing really has to be changed from Popa et al (2013)'s mOPE scheme. The client will take the value $x$ to insert, create $x'$, then encrypt $x'$ to obtain a ciphertext. Inserting this ciphertext will involve moving down the server's binary search tree. The insertion process is started by the client acquiring the root node. The client then checks to see if $x'$ is larger or smaller than the root's decrypted value. This informs the client as to the direction in which it should move. Once it arrives at a leaf node, the client will insert the new ciphertext in the correct direction.

@@

Querying, however, gets more difficult, and the next section will discuss how the client can query for some value $y$. 

@@

  To query for some value $y$, we must actually perform a range query, since any value $y$ will be encrypted with a plaintext $y + r$ where $r \in [0,1)$. This means, the encrypted plaintext of $y$ can therefore take in the range $[y, y + 1)$. Thus, to query for the value $y$, we must perform a range query on any values in the range $[y, y+1)$. \\ \\

@@

Now suppose $x$ is not in the subtree rooted at $l$. This means that $x$ must either be smaller than the left-most node in the subtree rooted at $l$ or greater than the right-most node in the subtree (assuming that no nodes are have repeated values which occurs with high probability because of the random bits $r$ appended to each plaintext). However, we know that the left-most node in the subtree rooted at $l$ must be smaller than $y$ and the right-most node in the subtree must be greater than $y+1$. This is true because both $y$ and $y+1$ are nodes in the subtree. \\ \\

@@

\end{proof} \\ \\This lemma instructs us on how to find all values $x \in [y, y+1)$. As we move up the binary search tree to parents of $y$, we check the right subtree of the parent looking for all values in $[y, y+1)$. We do this as we continue upwards for all successive parent nodes. For the node at $y+1$, we check the left subtree of all parents for values in $[y, y+1)$. We continue this process until the parent of the $y$ node meets the parent of the $y+1$ node. This process will find all values in $[y, y+1)$, because as we have shown, all values in the interval are inside the subtree. \\ \\

@@

We will show this by induction on the number of repeated elements in either $V_i$ or $W_i$. We shall first consider the base case when there are no repeated elements in $V_i$ or in $W_i$, so that all the elements in $V_i$ are distinct, as well as the elements in $W_i$. In this case, the sequence $v$ can be broken down into $\kappa$ sequences of all distinct elements. This has a clear bijection to a sequence of elements $v_{ij}$ of length $\kappa \lambda$, for $i \in \{1, 2, \ldots, \kappa\}$ and $j \in \{1, 2, \ldots, \lambda \}$. Winning this game then becomes equivalent to winning the IND-OCPA security game, since one can think of each $v_{ij}$ and $w_{ij}$ pair as a set of values that the adversary provides the client in each round $i$. Popa et al (2013) shows that their scheme, which is equivalent to the Distribution Confidentiality Scheme when there are no reptitions, is IND-OCPA secure. This means that the DCS scheme is IND-CPDA secure when there are no repeated elements. \\ \\

@@

Consider a simple tie-breaking rule on top of Popa et al (2013)'s mOPE scheme. When the client inserts a new ciphertext $c^*$ into the server, the client proceeds normally when the ciphertext $c_i$ returned by the server does not correspond to a plaintext $p_i$ which is equal to $p^* = Dec(SK, c^*)$. However, if $p_i = p^*$, then the client tells the server to move left or right with uniform probability. Note that in this scheme, we will use a randomized encryption scheme which is IND-CCA so that the server cannot tell if two plaintext values are the same. \\ \\

@@

This scheme is a simple extension of Popa et al (2013)'s scheme and handles repeated values. In RTOPE, the server cannot tell how many values are repeated, and what the distribution of repeated values looks like.

@@

Under RTOPE, there are multiple variations on querying for some value $x$. First is to query for just one ciphertext whose corresponding plaintext is equal to $x$. The second query type is to query for all ciphertexts whose corresponding plaintext is equal to $x$. \\ \\

@@

When there are no repeated elements, we can use the same argument we used in the DCS proof. We note that the sequence $\{V_1, V_2, \ldots, V_{\kappa} \}$ has a bijection to a sequence of elements $v_{ij}$ of length $\kappa \lambda$, for $i \in \{1, 2, \ldots, \kappa\}$ and $j \in \{1, 2, \ldots, \lambda \}$. Winning this game then becomes equivalent to winning the IND-OCPA security game, since one can think of each $v_{ij}$ and $w_{ij}$ pair as a set of values that the adversary provides the client in each round $i$. Popa et al (2013) shows that their scheme, which is equivalent to the Distribution Confidentiality Scheme when there are no reptitions, is IND-OCPA secure. Thus, the base case of no repetitions is secure. \\ \\

@@

  This proposition is immediately clear because one can always take the entire tree as the subtree. This works for the cipher text trees stored in both DCS and RTOPE.

@@

To confirm our theoretical analysis, we implemented both the DCS and RTOPE encryption schemes on top of a simplified version of CryptDB. First, we created a framework that mimicked the necessary features of CryptDB, including the client-server interaction and full insertion and query functionality. We then made a testing framework which allowed us to easily change the number of inputs, the number of repeats in the inputs, and the choice of whether to use DCS or RTOPE. The results are summarizes in the table below. \\ \\

@@

\item The standard deviation in OPE re-encodings varies significantly between the two schemes. This has to do with the resulting shape of the trees formed for the same inputs -- DCS is more likely to create an unbalanced tree, and thus the standard deviations tend to be higher than comparable RTOPE deviations.

@@

We proved that DCS and RTOPE are IND-CPDA secure and introduced the notion of structurally expectation equivalence. We also proved the important theorem that if two OPE schemes are structurally expectation equivalent, then one is IND-CPDA secure if and only if the other is IND-CPDA secure as well. We then showed that though DCS and RTOPE do satisfy this structurally expectation equivalence relationship, they are still IND-CPDA secure. \\ \\
We noted that though the schemes are not structurally expectation equivalent, the number of OPE re-encodings will converege to the same number, given identical inputs. Finally, this notion was further supported by our experimental results. We tested our implementation of DCS and RTOPE on a range of inputs, and found that the number of OPE re-encodings were of the same order.

>>>>

Order-preserving encryption schemes serve to generate ciphertexts that preserve the numerical ordering of the plaintexts. Such schemes provide an entity for storing encrypted data, such as a mail server or a server interacting with a database, and the ability to perform order comparison operations such as sorting, range query checks, and ranking, among other applications. The ideal security goal for an order-preserving encryption scheme, IND-OCPA, is for the scheme to reveal nothing except for the order of the plaintexts. \\ \\
For more than a decade, researchers have developed such schemes with varying degrees success. Often, thees schemes compromised security to preserve order, leaking plaintexts or other attributes of the data. Recently, the first order-preserving encryption scheme that satisfies IND-OCPA, mutable order-preserving encoding (mOPE), was proposed by Popa et al (2013). mOPE works by creating a balanced search tree containing all the encrypted data on a server, giving the server the ability to order the tree based on the plaintexts without learning any information regarding the plaintexts themselves. All encrypted values are represented by a node in the balanced search tree. Each node has an Order Preserving Encoding (OPE) which represents its path on the tree from the root to itself, allowing the server the ability to maintain the order of the plaintexts. \\ \\ 
One disadvantage of mOPE, however, is that multiple entries of the same plaintext value are maintainted in the same node in the tree with a counter. This means that the scheme can leak information about the distribution of plaintext values. In this paper, we present two schemes that address this issue: Distribution Confidentiality Scheme (DCS) and Random Tiebreaking Order Preservering Encryption (RTOPE). Further, we define a new security guarantee, Indistinguisability Under Chosen Plaintext Distribution Attack (IND-CPDA), which these schemes are intended to satisfy. \\ \\

@@

The mOPE scheme by Popa et al (2013) is the first scheme to achieve ideal IND-OCPA security, and our schemes are built upon mOPE. Prior work on range query schemes all leak information about the underlying plaintext. Shi et al (2007) designed an encryption scheme that leaks the decrypted data for values that fall within the range query, but preserves the privacy of all other values. We chose to build on schemes on existing OPE schemes, instead of range query schemes, to guarantee that security of underlying plaintext even within the range queries. \\ \\
Further, Popa et al (2013) provides a new security definition, same-time OPE security, which dictates that an adversary can learn the order only for values stored in the application at the same time, and is stronger than IND-OCPA. Our security guarantee of IND-CPDA, explained later in the paper, asks furthermore that the ciphertexts not leak the distribution of values. 

@@

In the basic version of the scheme, let us assume that we want to encrypt some plaintext value $x \in \mathrm{Z}$ and send it to the server like in the mOPE scheme. For DCS, there is an additional parameter $\rho$ which will govern the probability with which the server sees a repeated value. We will generate a random value $r \in [0,1)$ which contains $\rho$ bits, denoted $r_{\rho}$, and append it to the end of $x$. The resulting value $x' = x + r_{\rho}$ will be encrypted with DET. The mOPE scheme will work the same as before, but now acting on $x'$. \\ \\
Inserting in this scheme is simple, since nothing else has to be changed from the mOPE scheme. The client will take the value $x$ to insert, create $x'$, then encrypt $x'$ to obtain a ciphertext. Inserting this ciphertext will involve moving down the server's binary search tree. The insertion process is started by the client acquiring the root node. The client then checks to see if $x'$ is larger or smaller than the root's decrypted value. This informs the client as to the direction in which it should move. Once it arrives at a leaf node, the client will insert the new ciphertext in the correct direction.

@@

Querying, however, is more difficult, and the next section will discuss how the client can query for some value $y$. 

@@

  To query for some value $y$, we must actually perform a range query, since any value $y$ will be encrypted with a plaintext $y + r$ where $r \in [0,1)$. Therefore, the encrypted plaintext of $y$ can be in the range $[y, y + 1)$. Thus, to query for the value $y$, we must perform a range query on any values in the range $[y, y+1)$. \\ \\

@@

Now suppose $x$ is not in the subtree rooted at $l$. This means that $x$ must either be smaller than the left-most node in the subtree rooted at $l$ or greater than the right-most node in the subtree (assuming that no nodes are have repeated values, which occurs with high probability because of the random bits $r$ appended to each plaintext). However, we know that the left-most node in the subtree rooted at $l$ must be smaller than $y$ and the right-most node in the subtree must be greater than $y+1$. This is true because both $y$ and $y+1$ are nodes in the subtree. \\ \\

@@

\end{proof} \\ \\This lemma instructs us on how to find all values $x \in [y, y+1)$. As we move up the binary search tree to parents of $y$, we check the right subtree of the parent looking for all values in $[y, y+1)$. We do this as we continue upwards for all successive parent nodes. For the node at $y+1$, we check the left subtree of all parents for values in $[y, y+1)$. We continue this process until the parent of the $y$ node meets the parent of the $y+1$ node. This process will find and return all values in $[y, y+1)$, because as we have shown, all values in the interval are inside the subtree. \\ \\

@@

We will show this by induction on the number of repeated elements in either $V_i$ or $W_i$. We shall first consider the base case when there are no repeated elements in $V_i$ or in $W_i$, so that all the elements in $V_i$ are distinct, as well as the elements in $W_i$. In this case, the sequence $v$ can be broken down into $\kappa$ sequences of all distinct elements. This has a clear bijection to a sequence of elements $v_{ij}$ of length $\kappa \lambda$, for $i \in \{1, 2, \ldots, \kappa\}$ and $j \in \{1, 2, \ldots, \lambda \}$. Winning this game then becomes equivalent to winning the IND-OCPA security game, since one can think of each $v_{ij}$ and $w_{ij}$ pair as a set of values that the adversary provides the client in each round $i$. Popa et al (2013) show that their scheme, which is equivalent to the Distribution Confidentiality Scheme when there are no reptitions, is IND-OCPA secure. This means that the DCS scheme is IND-CPDA secure when there are no repeated elements. \\ \\

@@

Consider a simple tie-breaking rule on top of the mOPE scheme. When the client inserts a new ciphertext $c^*$ into the server, the client proceeds normally when the ciphertext $c_i$ returned by the server does not correspond to a plaintext $p_i$ which is equal to $p^* = Dec(SK, c^*)$. However, if $p_i = p^*$, then the client tells the server to move left or right with uniform probability. Note that in this scheme, we will use a randomized encryption scheme which is IND-CCA so that the server cannot tell if two plaintext values are the same. \\ \\

@@

This scheme is a simple extension of the mOPE scheme and handles repeated values. In RTOPE, the server cannot tell how many values are repeated, nor what distribution of repeated values looks like.

@@

Under RTOPE, there are multiple variations of querying for some value $x$. First is to query for just one ciphertext whose corresponding plaintext is equal to $x$. The second query type is to query for all ciphertexts whose corresponding plaintext is equal to $x$. \\ \\

@@

When there are no repeated elements, we can use the same argument we used in the DCS proof. We note that the sequence $\{V_1, V_2, \ldots, V_{\kappa} \}$ has a bijection to a sequence of elements $v_{ij}$ of length $\kappa \lambda$, for $i \in \{1, 2, \ldots, \kappa\}$ and $j \in \{1, 2, \ldots, \lambda \}$. Winning this game then becomes equivalent to winning the IND-OCPA security game, since one can think of each $v_{ij}$ and $w_{ij}$ pair as a set of values that the adversary provides the client in each round $i$. Popa et al (2013) shows that their scheme, which is equivalent to the Distribution Confidentiality Scheme when there are no repetitions, is IND-OCPA secure. Thus, the base case of no repetitions is secure. \\ \\

@@

  This proposition is immediately clear because one can always take the entire tree as the subtree. This works for the ciphertext trees stored in both DCS and RTOPE.

@@

To confirm our theoretical analysis, we implemented both the DCS and RTOPE encryption schemes on top of a simplified version of CryptDB. First, we created a framework that mimicked the necessary features of CryptDB, including the client-server interaction and full insertion and query functionality. We then made a testing framework which allowed us to easily change the number of inputs, the number of repeats in the inputs, and the choice of whether to use DCS or RTOPE. The results are summarized in the table below. \\ \\

@@

\item The standard deviation in OPE re-encodings varies significantly between the two schemes. This has to do with the resulting shape of the trees formed for the same inputs - DCS is more likely to create an unbalanced tree, and thus the standard deviations tend to be higher than comparable RTOPE deviations.

@@

We proved that DCS and RTOPE are IND-CPDA secure and introduced the notion of structurally expectation equivalence. We also proved the important theorem that if two OPE schemes are structurally expectation equivalent, then one is IND-CPDA secure if and only if the other is IND-CPDA secure as well. We then showed that though DCS and RTOPE do not satisfy this structurally expectation equivalence relationship, they are both IND-CPDA secure. \\ \\
We also noted that though the schemes are not structurally expectation equivalent, the number of OPE re-encodings will converege to the same number, given identical inputs. Finally, this notion was further supported by our experimental results. We tested our implementation of DCS and RTOPE on a range of inputs, and found that the number of OPE re-encodings were of the same order.
